---
title: "Class 22 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$




## Overview: Multiple linear regression continued



 * Comparing models with R^2, adjusted R^2, AIC and BIC
 * Cross-validation
 * Plotting using grammar of graphics with ggplot
 * Bonus ggplot features






$\\$







## Part 1: Comparing models with R^2, adjusted R^2, AIC, BIC and cross-validation


As we discussed, we can compare models using several measures including $R^2$, $R^2_{adj}$, $AIC$, $BIC$ and cross-validation. Let's try this out on the baseball data.




$\\$





#### Part 1.1: Comparing models with R^2, adjusted R^2, AIC, BIC


Let's compare our predictions of runs in baseball using the model with the original variables and the model that also includes the variables that includes statistics normalized by at bats.




```{r}


library(Lahman)
library(dplyr)



# reduce the data frame to the main batting statistics of interest
team_batting2 <- select(Teams, yearID, teamID, G,  W,  L, R,  AB,  H,  X2B, X3B,  HR,  BB,  SO, SB, CS,  HBP, SF) %>%
  mutate(X1B = H - (X2B + X3B + HR),
         X1Bn = X1B/AB, 
         X2Bn = X2B/AB, 
         X3Bn = X3B/AB, 
         XHRn = HR/AB, 
         XBBn = BB/AB)                        




fit1 <- lm(R ~ X1B + X2B + X3B + HR + BB + AB, data = team_batting2)
fit2 <- lm(R ~ X1B + X2B + X3B + HR + BB + X1Bn + X2Bn + X3Bn + XHRn + XBBn, data = team_batting2)


# summarize the fits



cat('R^2 \n')




cat("\n Adjusted R^2 \n")




cat("\n AIC \n")



cat("\n BIC \n")




```






$\\$






#### Part 1.2: Cross-validation


Let's now compare the models using cross-validation. To keep things simple, we are just going to split the data once into a training set and a test set rather than do k-fold cross-validation. 

We will compare the models based on their mean squared prediction error (MSPE) where a smaller MSPE is better. 



```{r}

# create the training set and the test set
total_num_points <- dim(team_batting2)[1]
num_training_points <- floor(total_num_points/2)

training_data <- team_batting2[1:num_training_points, ]
test_data <- team_batting2[(num_training_points + 1):total_num_points, ]



# fit both models on the training data, and calculate the MSPE based on thier predictions on the test set

fit_cv_1 <- lm(R ~ X1B + X2B + X3B + HR + BB + AB, data = training_data)

# get the test predictions and the MSPE for the smaller model 



fit_cv_2 <- lm(R ~ X1B + X2B + X3B + HR + BB + X1Bn + X2Bn + X3Bn + XHRn + XBBn, data = training_data)


# get the test predictions and the MSPE for the larger model 



# print the MSPEs 




```







$\\$







## Part 2: Visualization with ggplot


We can use the ggplot2 package which is part of the tidyverse to create much nicer looking graphics than using base R graphics. The ggplot2 library is modeled on Leland Wilkinson's "grammar of graphics" which tries to create graphics as a combination of basic visual elements. 


In the exercises below, we will use the motor trends cars data set (mtcars) that comes with base R installation, and also the gapminder data. 





$\\$





#### Part 2.1: Scatter plots


```{r}


# install.packages("ggplot2")


library(ggplot2)


# base R


# ggplot - mapping in the geom



# ggplot - global mapping




# ggplot - shorter global mapping




```




$\\$





#### Part 2.2: Adding labels to plots


We can add labels to the plots using the xlab("label1") and ylab("label2") functions



```{r}





```





$\\$





#### Part 2.3: More aesthetic mappings


```{r}


# add color based on the transmission type (is automatic or not)



# it is better to treat the transmission type as a categorical variable



# can also try mappping transmission type to shape or size 



```



$\\$



#### Part 2.4: Attributes vs. Aesthetics


Setting an **aesthetic mapping** maps a variable to a **glyph property**

Setting an **attribute** set a glyph property to a **fixed value**



```{r}


# setting an aesthetic



# setting an attribute




```




$\\$



#### Part 2.5: Facets


Beyond comparing variables based on aesthetics you can compare categorical variables by splitting a plot into subplots (called facets) using facet_wrap()




```{r}


# separate subplots for the two transmission types




# One can also do facets in two dimensions





```




$\\$




#### Part 2.6: Overplotting


Sometimes points overlap making it hard to estimate the number of points at a particular range of values. 

We can control the transparency of points by changing their alpha values.




```{r}

library(gapminder)


# a lot of overplotting




# chaning the transparency levels makes it a bit easir to see how many points are at a given x, y location






```





$\\$





#### Part 2.7a: geoms


We can use different geoms to create other types of plots.



```{r}

# Let's create a plot that shows the GDP in the United States as a function of the year using the geom geom_line()




# Let's plot a histogram of the weights of cars




# Let's create a boxplot of the weights of cars




# Let's create a side-by-side boxplot of the weights of cars depending on the number of cylinders the engine has





```





$\\$





#### Part 2.7b: geoms continued

Violin and Joy plots are other ways to view distributions of data



```{r}

# violin plot 


# library("ggridges")


# joy plot




```





$\\$





#### Part 2.8: geoms continued


We can also have multiple geom layers on a single graph by using the + symbol
E.g   ggplot(…) + geom_type1() + geom_type2()



```{r}

# Create a scatter plot of miles per gallon as a function of weight and then add a smoothed line using geom_smooth()



```





$\\$





#### Part 2.9: Themes


We can also use different types to change the appearance of our plot

```{r}


# Add theme_classic() to our plot



```




$\\$





## Part 3: Bonus features of ggplot


Below are some extra packages that can extend the functionality of ggplot




$\\$






#### Part 3.1: Additional geoms

There are additional packages that add more geoms




```{r}


# # install.packages("emoGG")

# # devtools::install_github("dill/emoGG")


# library(emoGG)




```




$\\$



#### Part 3.2: Adding animation


We can create animated images (gifs) using the gganimate package. To learn how to use this package see the [gganimate github website](https://github.com/thomasp85/gganimate)



```{r}

# # install.packages("gganimate")

# library(gganimate)



# example from:  https://github.com/thomasp85/gganimate

# run this from the console

# ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
#   geom_point(alpha = 0.7, show.legend = FALSE) +
#   scale_colour_manual(values = country_colors) +
#   scale_size(range = c(2, 12)) +
#   scale_x_log10() +
#   facet_wrap(~continent) +
#   # Here comes the gganimate specific bits
#   labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
#   transition_time(year) +
#   ease_aes('linear')


```




$\\$





#### Part 3.2: Interactivity


We can make our plots interactive using the plotly() library. To learn more about plotly [vist their website](https://plot.ly/)




```{r}


# # install.packages("plotly")

# library(plotly)

# also better run from the console or need to knit to html

# p <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, 
# 	size = pop, col = continent, frame = year)) +  
# 	geom_point() +  
# 	scale_x_log10()
# 
# ggplotly(p)



```








