---
title: "Class 24 notes and code"
output:
  pdf_document: 
   fig_width: 7
   fig_height: 5
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)

library(dplyr)


library(ggplot2)

#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```






$\\$






## Overview: Multiple linear regression continued

 * Bonus ggplot features
 * Joining data frames
 * Spatial mapping
 * Reshaping data






$\\$









## Part 1: Bonus features of ggplot


Below are some extra packages that can extend the functionality of ggplot




$\\$



#### Part 1.0: Renaming factors


Last class the question came up about renaming factors. We can use the recode() function in the dplyr package to do this.


For more information see the this paper by [McNamara and Horton](https://peerj.com/preprints/3163/). For other useuful function related to factors see the [forcats package](https://forcats.tidyverse.org/)



```{r}



# mutating the names of a factor to be more meaningful using dplyr recode()




```










$\\$





#### Part 1.1: Additional geoms

There are additional packages that add more geoms




```{r}


# # install.packages("emoGG")

library(emoGG)


# plot car mpg as a function of weight using the geom: geom_emoji("1f697")




```




$\\$



#### Part 1.2: Adding animation


We can create animated images (gifs) using the gganimate package. To learn how to use this package see the [gganimate github website](https://github.com/thomasp85/gganimate)



```{r}

# # install.packages("gganimate")

# library(gganimate)


# example from:  https://github.com/thomasp85/gganimate

# run this from the console

# ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
#   geom_point(alpha = 0.7, show.legend = FALSE) +
#   scale_colour_manual(values = country_colors) +
#   scale_size(range = c(2, 12)) +
#   scale_x_log10() +
#   facet_wrap(~continent) +
#   # Here comes the gganimate specific bits
#   labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
#   transition_time(year) +
#   ease_aes('linear')




```




$\\$





#### Part 1.2: Interactivity


We can make our plots interactive using the plotly() library. To learn more about plotly [vist their website](https://plot.ly/)




```{r}


# # install.packages("plotly")

# library(plotly)

# also better run from the console or need to knit to html

# p <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, 
# 	size = pop, col = continent, frame = year)) +  
# 	geom_point() +  
# 	scale_x_log10()
# 
# ggplotly(p)



```



$\\$






## Part 2: Joining data frames


Often data of intertest is spread across multiple data frames that need to be joined together into a single data frame for further analyses. We will explore how to do this using dplyr. 


Let's look at a very simple data set to explore joining data frames. 


```{r}



# download_class_data('x_and_y.Rda')

#load('x_and_y.Rda')





```


$\\$



#### Part 2.1: Left join

Left joins keep all rows in the left table.  

Data from right table added when there is the key matches, otherwise NA as added. 



```{r}



```



$\\$



#### Part 2.2: Right join


Right joins keep all rows in the right table.  

Data from left table added when there is the key matches, otherwise NA as added. 




```{r}



```



$\\$



#### Part 2.3: Inner join


Inner joins only keep rows in which there are matches between the keys in both tables



```{r}




```



$\\$



#### Part 2.4: Full join

Full joins keep all rows in both table.  
NAs are added where there are no matches



```{r}



```



$\\$




#### Part 2.5a: Duplicate keys


Duplicate keys are useful if there is a one-to-many relationship (duplicates are usually in the left table). 

Let's look at two other tables that have duplicate keys



```{r}



```




$\\$





#### Part 2.5b: Duplicate keys


If both tables have duplicate keys you get all possible combinations (Cartesian product). This is almost always an error! Always check the output dimension after you join a table because even if there is not a syntax error you might not get the table you are expecting!



```{r}


# initial left data frame only has 3 rows


# left join when both the left and right tables have duplicate keys


# output now has more rows than the initial table



```




$\\$




#### Part 2.5c: Duplicate keys



To deal with duplicate keys in both tables, we can join the tables **using multiple keys** in order to make sure that each row is uniquely specified


```{r}


# initial left data frame only has 3 rows


# join with mulitple keys



```




$\\$ 




## Part 3: Creating cloropleth maps



To practice creating cloropleth maps, let's plot data some demographic data which dates have growing numbers of younger people. 


From more fun data sets to map, check out [howmuch.net](https://howmuch.net/sources/income-inequality-by-state)



$\\$






#### Part 3.1: Get map coordinates



```{r}


# Get the map of the States


# Plot a state



```




$\\$



#### Part 3.2: Combine demographic data with the map coordinates



```{r}


# download the data
# download_class_data("state_population.Rda")


# load the data
#load("state_population.Rda")


# get data from only 2013


# join state_pop_2013 and states_map data frames 


# order the data



```





$\\$



#### Part 3.3: Plot the map



```{r}





```



**Question:**  What is wrong with this figure?  [Hint](https://xkcd.com/1138/)





$\\$






#### Part 3.4: Normalizing data



```{r}






```





$\\$







## Part 4: Reshaping data


Sometimes it is more convenient to have data in a data frame that is in either a longer or a wider format. The tidyr package as two functions pivot_longer() and pivot_wider() which are useful for this. 


Let us explore data that has the weather in different cities to understand why these functions are useful. 



```{r}


# use the tidyr package to reshape data
library(tidyr)


# download the data
# download.file('https://raw.githubusercontent.com/emeyers/SDS230_F19/master/class_data/histWeather.csv', 'histWeather.csv')


# load the data into R


# convert variables to the appropriate types




```





$\\$





#### Part 4.1: Creating longer data


Suppose we want to plot each of the weather measurements as a function of the data on a separate subplot. To do this it is convenient to first convert the data to a wide format and then we can use ggplot's facet_wrap() to plot each measurement. 


```{r}


# create longer data for New Haven only


# plot the data



```





$\\$





#### Part 4.2: Creating wider data


Suppose we want to compare the minimum temperature in New York City to the minimum temperature in New Haven for each date of the year. To do this it would be useful to the minimum temperature of New York City and of New Haven in seperate columns. We can then subtract the values in these columns to get the temperature difference. 



```{r}


# just get the minimum temperature for New York City and New Haven




# make a wider data frame that has separate columns for New York City and New Haven



# Let's plot how much warmer New York City's minimum temperature is on each date





```


















