---
title: "Class 10 notes and code"
output:
  html_document: default
  pdf_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$




## Overview

 * Review permutation tests for two means in R
 * Permutation tests for more than 2 means
 * Permutation tests for more than 2 means in R
 * Review z-scores and correlation
 * Permutation tests for correlation



$\\$









## Part 1: Permutation tests for comparing two sample means in R


**Question:** Does eating late at night make you put on more weight compared to whether you ate the same amount of food during the day? 


To test this idea researchers, [Fonken et al, 2010](https://www.ncbi.nlm.nih.gov/pubmed/20937863), examined this question in mice. 

Mice were randomly divided into 2 groups:

 * **Dark condition**:  8 mice were given 8 hours of darkness at night when they couldn’t eat
 * **Light condition**: 9 were constantly exposed to light for 24 hours so they could always eat


The data from the mice is below. Let's use a permutation test to see if there is any difference between the amount of weight the mice gained between these two conditions. 





##### Step 0: Plot the data


```{r mice_plot}

# run this once
# get the mice data
# source('class_functions.R')
# download_class_data("mice.Rda")

# load the data
load("mice.Rda")



# plot the data as a boxplot



# plot the data as a strip chart by:



# a) creating a list of the data 


# b) creating a stripchart   








```




##### Step 1: State the null and alternative hypotheses 


$H_0: $
$H_A: $


$\alpha = $





$\\$





##### Step 2: Calculate the observed statistic

```{r pill_obs_stat}




```



$\\$





##### Step 3: Create the null distribution



```{r null_dist_mice}


# combine the data from the dark and light groups together


# use a for loop to create shuffled dark and light groups and shuffled statistics 
null_distribution <-  NULL
for (i in 1:10000) {

       # shuffle data

        # create fake dark and light groups


        # save the statistic of interest

}






```



$\\$





##### Step 4: Calculate a p-value


```{r p_value_mice}


# plot the null distribution again with a red line a the value of the observed statistic




# calculate the p-value




```




$\\$






##### Step 5: Make a decision








$\\$





[Give a mouse a cookie](https://hamstersworld.weebly.com/uploads/2/0/6/1/2061190/8024088.jpg)


[science has gone too far](https://i.imgflip.com/9vr.jpg)








$\\$






## Part 2: Permutation tests for more than two sample means


A similar test can be run comparing more than 2 means and for seeing if there is a meaningful correlation in the data...






$\\$








## Part 3: Permutation tests for more than two sample means in R 



A group of Hope College students wanted to see if there was an association between a student’s major and the time it takes to complete a small Sudoku-like puzzle

They grouped majors into four categories
 * Applied science (as)
 * Natural science (ns)
 * Social science  (ss)
 * Arts/humanities (ah)





$\\$





##### Step 0: Plot the data


```{r sudoku_plot}


# get the data
sudoku_data <- read.table('http://www.math.hope.edu/isi/data/chap9/MajorPuzzle.txt', header = TRUE)



# create a boxplot of the data





```




$\\$





##### Step 1: State the null and alternative hypotheses 


$H_0: $
$H_A: $ 


$\alpha = $



$\\$








##### Step 2: Calculate the observed statistic




```{r sudoku_obs_stat}


# we can use the by() function to get the means separately for each group
# (we will learn of a better way to do this later in the semester)
group_means <- as.vector(by(sudoku_data$time, sudoku_data$major, mean))


# a function to get the MAD statistic
get_MAD_stat <- function(group_means) {
  
  total <- 0
  num_differences <- 1
  
  for (iGroup1 in 1:(length(group_means) - 1)) {
    
    for (iGroup2 in (iGroup1 + 1):(length(group_means))){
      
      total <- total + abs(group_means[iGroup1] - group_means[iGroup2])
      num_differences <- num_differences + 1
      
    }
  
  }
  
  total/num_differences
  
}  # end of the function 



# apply this function to our data to get our observed statistic 



  

```






$\\$




##### Step 3: Create the null distribution


```{r sudoku_null_dist}


# create the null distribution here using a for loop and the get_MAD_stat() function 
null_dist <- NULL
for (i in 1:10000){
  
  
}


# plot the null distribution as a histogram



```





$\\$




##### Step 4: Calculate a p-value


```{r sudoku_p_value}



# plot the null distribution again with a line at the observed statistic



# calculate the p-value



```





$\\$







##### Step 5: Make a decision







$\\$







## Part 4: Review z-scores and correlation


Let's review correlation and how we could create a permutation test to see if there is a statistically significant correlation between two quantitative variables. 





$\\$






## Part 5: Permutation tests for correlation





$\\$





##### Step 0: Plot the data


```{r cereal_plot}


# get the data
# source('class_functions.R')
# download_class_data("cereal.Rda")

# load the data
load("cereal.Rda")


# create a scatter plot of carbs as a function of calories





```





$\\$







##### Step 1: State the null and alternative hypotheses 


$H_0: $
$H_A: $ 


$\alpha = $



$\\$








##### Step 2: Calculate the observed statistic




```{r cor_obs_stat}



# calculate the correlation as our observed statistic 

  

```






$\\$




##### Step 3: Create the null distribution




```{r cor_null_dist}


# one null distribution statistic 
cor(sample(cereal$Calories), cereal$Carbs)


# create the full null distribution 




```





$\\$




##### Step 4: Calculate a p-value


```{r cor_p_value}



# plot the null distribution again with a line at the observed statistic



# calculate the p-value





```


$\\$







##### Step 5: Make a decision








$\\$






## Homework 5: Was the 1969 Vietnam Draft completely random? 







$\\$












