---
title: "Class 23 notes and code"
output:
  pdf_document: 
   fig_width: 7
   fig_height: 5
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)

library(dplyr)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```






$\\$






## Overview: Multiple linear regression continued



 * Plotting using grammar of graphics with ggplot
 * Bonus ggplot features
 * Joining data frames






$\\$








## Part 1: Visualization with ggplot


We can use the ggplot2 package which is part of the tidyverse to create much nicer looking graphics than using base R graphics. The ggplot2 library is modeled on Leland Wilkinson's "grammar of graphics" which tries to create graphics as a combination of basic visual elements. 


In the exercises below, we will use the motor trends cars data set (mtcars) that comes with base R installation, and also the gapminder data. 





$\\$





#### Part 1.1: Scatter plots


```{r}


# install.packages("ggplot2")


library(ggplot2)


# base R


# ggplot - mapping in the geom



# ggplot - global mapping
 


# ggplot - shorter global mapping




```




$\\$





#### Part 1.2: Adding labels to plots


We can add labels to the plots using the xlab("label1") and ylab("label2") functions



```{r}






```





$\\$





#### Part 1.3: More aesthetic mappings


```{r}


# add color based on the transmission type (is automatic or not)




# it is better to treat the transmission type as a categorical variable




# can also try mappping transmission type to shape or size 




```



$\\$



#### Part 1.4: Attributes vs. Aesthetics


Setting an **aesthetic mapping** maps a variable to a **glyph property**

Setting an **attribute** set a glyph property to a **fixed value**



```{r}


# setting a aesthetic



# setting an attribute




```




$\\$



#### Part 1.5: Facets


Beyond comparing variables based on aesthetics you can compare categorical variables by splitting a plot into subplots (called facets) using facet_wrap()




```{r}


# separate subplots for the two transmission types




# One can also do facets in two dimensions





```




$\\$




#### Part 1.6: Overplotting


Sometimes points overlap making it hard to estimate the number of points at a particular range of values. 

We can control the transparency of points by changing their alpha values.




```{r}

library(gapminder)


# a lot of overplotting



# changing the transparency levels makes it a bit easir to see how many points are at a given x, y location




```





$\\$




#### Part 1.7: Changing scales


We can transform the scales of continuous data too


```{r}

# changing the scale on the x-axis



```






$\\$






#### Part 1.8: geoms


We can use different geoms to create other types of plots.



```{r}

# Let's create a plot that shows the GDP in the United States as a function of the year using the geom geom_line()




# Let's plot a histogram of the weights of cars



# Let's create a boxplot of the weights of cars




# Let's create a side-by-side boxplot of the weights of cars depending on the number of cylinders the engine has





```





$\\$





#### Part 1.9: geoms continued

Violin and Joy plots are other ways to view distributions of data



```{r}

# violin plot 




# library("ggridges")

# joy plot





```





$\\$





#### Part 1.10: geoms continued continued


We can also have multiple geom layers on a single graph by using the + symbol
E.g   ggplot(…) + geom_type1() + geom_type2()



```{r}

# Create a scatter plot of miles per gallon as a function of weight and then add a smoothed line using geom_smooth()



```





$\\$





#### Part 1.10: Themes


We can also use different types to change the appearance of our plot

```{r}


# Add theme_classic() to our plot




```




$\\$




#### Part 1.11: Example - plotting the IPED faculty salary data


```{r}


# source('../../../SDS230_F19/class_functions/class_functions.R')
# download_class_data("IPED_salaries_2016.rda")
# load("IPED_salaries_2016.rda")







```













## Part 2: Bonus features of ggplot


Below are some extra packages that can extend the functionality of ggplot




$\\$






#### Part 2.1: Additional geoms

There are additional packages that add more geoms




```{r}


# # install.packages("emoGG")

# library(emoGG)




```




$\\$



#### Part 2.2: Adding animation


We can create animated images (gifs) using the gganimate package. To learn how to use this package see the [gganimate github website](https://github.com/thomasp85/gganimate)



```{r}

# # install.packages("gganimate")

# library(gganimate)


# example from:  https://github.com/thomasp85/gganimate

# run this from the console

# ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
#   geom_point(alpha = 0.7, show.legend = FALSE) +
#   scale_colour_manual(values = country_colors) +
#   scale_size(range = c(2, 12)) +
#   scale_x_log10() +
#   facet_wrap(~continent) +
#   # Here comes the gganimate specific bits
#   labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
#   transition_time(year) +
#   ease_aes('linear')




```




$\\$





#### Part 2.2: Interactivity


We can make our plots interactive using the plotly() library. To learn more about plotly [vist their website](https://plot.ly/)




```{r}


# # install.packages("plotly")

# library(plotly)

# also better run from the console or need to knit to html

# p <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, 
# 	size = pop, col = continent, frame = year)) +  
# 	geom_point() +  
# 	scale_x_log10()
# 
# ggplotly(p)



```



$\\$






## Part 3: Joining data frames


Often data of intertest is spread across multiple data frames that need to be joined together into a single data frame for further analyses. We will explore how to do this using dplyr. 


Let's look at a very simple data set to explore joining data frames. 


```{r}



#  download_class_data('x_and_y.Rda')
# load('x_and_y.Rda')



```


$\\$



#### Part 3.1: Left join

Left joins keep all rows in the left table.  

Data from right table added when there is the key matches, otherwise NA as added. 



```{r}



```



$\\$



#### Part 3.2: Right join


Right joins keep all rows in the right table.  

Data from left table added when there is the key matches, otherwise NA as added. 




```{r}




```



$\\$



#### Part 3.3: Inner join


Inner joins only keep rows in which there are matches between the keys in both tables



```{r}





```



$\\$



#### Part 3.4: Full join

Full joins keep all rows in both table.  
NAs are added where there are no matches



```{r}




```



$\\$




#### Part 3.5a: Duplicate keys


Duplicate keys are useful if there is a one-to-many relationship (duplicates are usually in the left table). 

Let's look at two other tables that have duplicate keys



```{r}




```




$\\$





#### Part 3.5b: Duplicate keys


If both tables have duplicate keys you get all possible combinations (Cartesian product). This is almost always an error! Always check the output dimension after you join a table because even if there is not a syntax error you might not get the table you are expecting!



```{r}

# initial left data frame only has 3 rows



# left join when both the left and right tables have duplicate keys


# output now has more rows than the initial table



```




$\\$




#### Part 3.5c: Duplicate keys



To deal with duplicate keys in both tables, we can join the tables **using multiple keys** in order to make sure that each row is uniquely specified


```{r}


# initial left data frame only has 3 rows




# output now only has 3 rows




```




$\\$ 





