---
title: "Class 18 notes and code"
output:
  pdf_document: default
  html_document: default
---





$\\$





```{r setup, include=FALSE}

# install.packages("latex2exp")

library(latex2exp)


#options(scipen=999)


knitr::opts_chunk$set(echo = TRUE)

set.seed(123)

```



$\\$




## Overview

 * Additional topics in simple linear regression: 
    * Confidence and prediction intervals, leverage, standarized and studentized residuals, Cook's distance
 
 * Multiple linear regression
 




$\\$






## Part 1: Additional statistics for evaluating simple linear regression fits


Let us use the data from the 2000 Florida election to look at additional metrics for evaluating simple linear regression fits. 






$\\$








#### Part 1.2: Loading the data and fitting the model


Let's start by loading the data and fitting a linear regression model predicting the number of 



```{r }


#download.file('https://raw.githubusercontent.com/emeyers/SDS230_F19/master/class_data/florida_vote_data_2000.Rda', 'florida_vote_data_2000.Rda')


load('florida_vote_data_2000.Rda')


# arrange the data based on the number of votes Bush received


# fit the linear regression model





```







$\\$








#### Part 1.2: Confidence intervals and prediction intervals




```{r}


# confidence intervals for the betas




# confidence interval for the regression line mu_y



# prediction interval for the regression line



# plot both confidence interval and the prediction interval




# plot confidence interval





# plot prediction interval



```







$\\$







#### Part 1.3: Leverage and Cook's distance




```{r}



# get the h_i "hat values"



# plot the 20 largest hat values



# plot Cook's distances and plot them



# plot residuals as a function of the hat values




# use the base R function to plot the standardized residuals with the hat values 
#  along with contours that contain constant Cook's distance values






```






$\\$












## Part 2: Multiple linear regression using baseball data


$\\$




#### Part 2.1: Creating a few commonly used baseball statistics


Let's explore which statistic has the highest correlation with the number of runs a team has scored. 

We can start by creating a few commonly used statistics/metrics for quantifying hitting ability by running the code below. 


```{r}


# install.packages('Lahman')

library(Lahman)
library(dplyr)


# View(Teams)



# reduce the data frame to the main batting statistics of interest
team_batting <- select(Teams, yearID, teamID, G,  W,  L, R,  AB,  H,  X2B, X3B,  HR,  BB,  SO, SB, CS,  HBP, SF)


# get commonly used composite statistics 
team_batting <- mutate(team_batting,
                       X1B = H - (X2B + X3B + HR), 
                       BA = H/AB, 
                       OBP = (H + BB)/(AB + BB),
                       SlugPct = (X1B + 2 * X2B + 3 * X3B + 4 * HR)/AB,
                       OPS = OBP + SlugPct)


# only use teams that have played the 162 games in a season (the number of games currently in a MLB season)
team_batting <- filter(team_batting, G == 162)

  

```





$\\$






#### Part 2.2: Which statistics are most correlated with runs



Let's look at the *correlation* between runs a team scores (R) and the following statistics:

 * HR:  home runs
 * BA:  batting average
 * SlugPct: slugging percentage
 * OBP:  on base percentage



```{r}




# correlate runs with HR



# correlate runs with BA



# correlate runs with SlugPct



# correlate runs with OBP






```




$\\$





#### Part 2.3: Visualizing the relationships



Let's visualize the correlation between the main variables of interest: R, HR, OBP, and SlugPct.



```{r}




```







$\\$



#### Part 2.4: Fitting a regression model to come up with a better metric


Let's now use the lm() function to fit a linear regression model that can potentially do a better job at predicting runs.

We can also evaluate which variables seem to be useful for predicting runs.




```{r}


# fit the multiple regression model



# examine the coefficients



# look at the model fit statistics



# extract the r^2 value



```



**Questions:**  Do the values for the coefficients make sense? 







$\\$







#### Part 2.5: Can you come up with a regression model that has an even higher correlation with runs? 


Add new variables to the data frame that contain additional statistics (e.g., HR^2, etc) and see if you can come up with an even higher correlation value with runs. 


Do you think your new model will be better at predicting the number of runs a team scores in the upcoming 2020 baseball season? 



```{r}




```






















